---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
# set the filename and load it into a data frame
datafile <- "activity/activity.csv"
setwd("C:/Users/david/datasciencecoursera/Reproducible Research/Week2/Project1/RepData_PeerAssessment1")
df <- read.csv(datafile)

```


## What is mean total number of steps taken per day?
Calculate the total steps by day

```{r}
# total the steps by date
totalSteps <- aggregate(steps ~ date, df, sum)
head(totalSteps)
```

Then make a histogram of the result

```{r}
hist(totalSteps$steps, main = "Histogram of Daily Steps",
     border = "dark blue", col = "light blue", xlab="Total Steps/day")
```

The mean of the daily steps

```{r}
# calculate the mean of the daily steps
mean(totalSteps$steps)
```

The median of the daily steps

```{r}
# and calculate the median number of steps
median(totalSteps$steps)
```

## What is the average daily activity pattern?
Make a time-series plot
First, let's determine the mean for each interval
```{r}
meanInterval <- aggregate( steps ~ interval, df, mean, na.action="na.exclude")
head(meanInterval)
```

Now do a time series plot

```{r}
plot( meanInterval$steps, meanInterval$interval, type="l")
```

Determine which 5 minute interval has the maximum number of steps

```{r}
# find the maximum number of steps in any interval
max(meanInterval$steps)
# what is the index of the maximum number of steps
which.max(meanInterval$steps)
# which interval is it
meanInterval[which.max(meanInterval$steps),1]
```


## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(df$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Are there differences in activity patterns between weekdays and weekends?

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?





## Are there differences in activity patterns between weekdays and weekends?
